% Section 1 -- Introduction. Flowing prose (RLCR/DCPO/RAG register): technical opening, moderate
% connected sentences, no signpost filler, no defensive framing, no em-dashes. Contributions as a
% short enumerate (the one \paragraph the RAG paper uses).
\section{Introduction}
\label{sec:intro}

\input{sections/F1_principle.tex}

Reinforcement learning with verifiable rewards (RLVR) post-trains a language model against a reward
computed from observed outcomes \citep{grpo, deepseek_r1}. Calibrated probabilistic prediction is a
natural target for it: a proper scoring rule such as the Brier score is computable from outcomes alone
and is minimized in expectation by the true probability \citep{brier1950, gneiting2007}, so optimizing
it should produce forecasts whose probabilities match observed frequencies.

In practice, reinforcement learning does the opposite, degrading calibration and leaving models
overconfident, whether the reward is human feedback \citep{leng2024} or a verifiable correctness signal
\citep{dcpo, uncalibrated_reasoning}. The work that corrects this addresses epistemic uncertainty, where
the model answers a question that has a correct answer and the goal is a calibrated confidence that the
answer is right \citep{rlcr, dcpo}.

A second kind of uncertainty is left untreated. It is aleatoric: the output is itself a probability, and
the label is a single realized outcome of a stochastic event, with no answer that can be called correct.
NFL in-game win probability is a clear instance. At any point in a game the forecaster states the
probability that the team in possession wins; the realized result is one Bernoulli draw from a rate that
no model observes directly; and the betting market provides a strong, independent estimate of that rate.
Whether RLVR can train a calibrated forecaster in this regime, and what makes it fail, has not been
studied.

Reinforcement learning against the per-play Brier score decalibrates such a forecaster through two
mechanisms. The label for a play is a single realized outcome, one draw from the rate being estimated,
so a reward scored against it has high variance and pulls the policy toward whichever result occurred.
When the model reasons in language before answering, a second problem appears: optimizing the final
probability rewrites the reasoning into incoherent arguments. We remove the variance by rewarding a
state-conditioned empirical win rate estimated from past outcomes, a verifiable and label-free target,
and we remove the corruption by keeping the policy gradient off the reasoning, through direct prediction
or a gradient mask over the answer tokens (\cref{fig:principle}). We train Qwen2.5-7B-Instruct without supervised fine-tuning
and evaluate on held-out seasons against the betting market, a frontier model, and the empirical-rate
table.

Trained by direct prediction, the model matches the market's calibration on held-out games, at an
expected calibration error of $0.029$ against the market's $0.027$, and it is better calibrated than a
zero-shot frontier model, with no human labels and no access to the market. On the Brier score it
matches that frontier model and a tabular estimator, the limit that the public game state allows; the
market's small remaining edge is the live in-game information none of them can see. Masking
the gradient instead of discarding the reasoning keeps the chain of thought faithful, lowering the rate
at which the stated probability does not follow from it from $22\%$ to $4\%$, at a small cost in
sharpness.

\paragraph{Contributions.}
\begin{enumerate}[leftmargin=1.4em,itemsep=2pt,topsep=2pt]
\item Our reward, a state-conditioned empirical win rate estimated from realized outcomes, is verifiable
and label-free, and replaces the single realized outcome, whose variance decalibrates per-play Brier
training; on held-out games it scores within $0.007$ Brier of the betting market
(\cref{sec:reward}).
\item Reinforcement learning decalibrates the forecaster through the gradient on the chain of thought,
not through the reward: with the rate target, training the full completion drives held-out ECE
from $0.19$ to $0.30$, whereas confining the gradient to the answer, by direct prediction or an
answer-span mask, holds it near the market's (\cref{sec:decoupling}).
\item With the chain of thought dropped, the model reaches the information ceiling of the public game
state: it matches the market's calibration ($0.029$ against $0.027$ ECE), is better calibrated than a
zero-shot frontier model, and
reaches the same Brier score as that frontier model and a tabular rate estimator, leaving the market's
edge as live in-game information (\cref{sec:results}).
\end{enumerate}
