% Section 8 -- Related work. DCPO/RLCR attribution register; long two-paper sentences split into
% one claim per sentence. Verified citations only.
\section{Related work}
\label{sec:related}

Reinforcement learning improves accuracy but degrades calibration, leaving models overconfident.
\citet{leng2024} attribute this, under human feedback, to a reward model that prefers confident answers.
\citet{dcpo} identify a conflict between the accuracy and calibration gradients under verifiable rewards,
and decouple the two with a masked-gradient update. \citet{uncalibrated_reasoning} attribute the
overconfidence on binary stochastic outcomes to the standard-deviation normalization in the
group-relative advantage, which they remove. These analyses take the realized outcome as the target.
With a denoised conditional-rate target we find instead that the normalization is necessary, and that the
overconfidence is driven by the gradient on the reasoning rather than by the advantage.

A second class of methods rewards a proper scoring rule on a verbalized confidence. \citet{rlcr} add a
Brier-score term to a correctness reward, \citet{rewarding_doubt} use the logarithmic score, and
\citet{band2024} reward a downstream reader's accuracy. These score a confidence in an answer that is
right or wrong. We score the forecast itself against a state-conditioned empirical rate, where the label
is a single stochastic outcome and correctness is undefined.

Language models are also trained and evaluated as event forecasters. \citet{halawi2024} approach
human-crowd accuracy through retrieval and aggregation, and \citet{pratt2024} find that prompting
strategies do not yield calibrated forecasts. \citet{turtel2025outcome} reward realized outcomes on
prediction-market and news questions, and \citet{turtel2025selftaught} rank self-generated forecasts
against outcomes. Both improve calibration while keeping the chain of thought and rewarding the outcome,
whereas we reward the conditional rate and mask the gradient off the reasoning. \citet{paleka2025} and
\citet{karger2024forecastbench} document temporal leakage in such evaluations, which our season-disjoint
test on resolved games avoids.

Our training is a choice of credit assignment within group-relative optimization \citep{grpo}. DAPO
\citep{dapo} and GPG \citep{gpg2026} remove the KL penalty and reference model, as we do. \citet{rpg2026}
analyze the instability of the KL estimator when the gradient concentrates on a few tokens.
\citet{wang2025entropy} restrict updates to high-entropy tokens, and \citet{archer2025, gtpo2025} treat
reasoning and answer tokens differently. Our answer-span mask restricts the gradient by token role.

Win-probability estimation for the NFL ranges from random forests \citep{lock2014} to the boosted models
of the nflverse pipeline \citep{yurko2018, nflfastr}, which we use as static baselines. \citet{brill2024}
show that play-by-play data bounds the accuracy of any such model, consistent with our reading of the
market's edge as structural. \citet{polson2015} treat the point spread as an implied volatility. The
betting market is known to predict outcomes more accurately than statistical models or individual bettors
\citep{boulier2003, levitt2004, franck2010, cox2021}. We convert market odds to probabilities following
\citet{strumbelj2014} and use them only as a reference.
